---
title: "Horn_bioclim_species_model"
author: "Malothro"
date: "2024-08-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(geodata)
library(raster)
library(ggmap)
library(dismo)
library(grid)
library(geosphere)
library(sf)
# Set Working Directory
setwd("C:\\Users\\dy0053tu\\Desktop\\r_data\\2_SDM+Data\\2_SDM Data")
```

## Load and prepare data
```{r}
horn <- read.csv("hornbill_my1.csv")
horn1 <- horn[,-1]  # Remove the first column
worldclim <- worldclim_global(var = "bio", res = 10, path = tempdir("C:\\Users\\dy0053tu\\Desktop\\r_data"))
```

## Define geographic extent

```{r}
ext <- extent(99, 105, 1.2, 6.7)
msia.worldclim <- crop(worldclim, ext)
```

## Define Bounding Box with Buffer

```{r}
buffer <- 1  # Adjust buffer distance as needed
h.extent <- extent(
  min(horn1$long) - buffer,
  max(horn1$long) + buffer,
  min(horn1$lat) - buffer,
  max(horn1$lat) + buffer
)
```

## Get Plots and Basemap

```{r}
stbbox <- c(left = xmin(h.extent), bottom = ymin(h.extent), right = xmax(h.extent), top = ymax(h.extent))
register_stadiamaps(key = "fd34d69d-d55b-4da7-af0b-bd6e38a855ab")
h.map <- get_stadiamap(bbox = stbbox, maptype = 'stamen_terrain', zoom = 5)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
```{r}
plot(h.map)
points(mercator(horn1[, c('long', 'lat')]), pch = 16, col = 'red')
h.map_raster <- rasterGrob(h.map, width = unit(1, "npc"), height = unit(1, "npc"))
horn1_sf <- st_as_sf(horn1, coords = c("long", "lat"), crs = 4326)
ggplot() +
  annotation_custom(h.map_raster, xmin = stbbox["left"], xmax = stbbox["right"], ymin = stbbox["bottom"], ymax = stbbox["top"]) +
  geom_sf(data = horn1_sf, color = "red", size = 3) +
  coord_sf(xlim = c(stbbox["left"], stbbox["right"]), ylim = c(stbbox["bottom"], stbbox["top"]), expand = FALSE) +
  theme_void()
```


```{r}
# Convert SpatRaster to RasterStack
msia.worldclim_raster <- raster::stack(msia.worldclim)

# Convert horn1 to a Matrix of Coordinates
coords_matrix <- as.matrix(horn1[c("long", "lat")])
```

```{r}
# Create Bioclim model
horn.d <- bioclim(msia.worldclim_raster, coords_matrix)

```

```{r}
# Plot Response Curves
par(mar = c(2,2,2,2), mfrow = c(4, 4))
response(horn.d)
par(mfrow = c(1, 1))
```

## Create Bioclim Model and Predict
```{r}
horn.d <- bioclim(msia.worldclim_raster, coords_matrix)
horn.d.pred <- predict(object = horn.d, x = msia.worldclim_raster)

# Plot Predictions
plot(horn.d.pred, main = 'SDM Predictions Using Climate Layers')
points(horn1[, c('long', 'lat')], pch = 16, cex = 0.25)
```

### evaluate model performance
### background data (pseudo-absences) needed for this
### determine if the model can differentiate bw the habitat & the background
```{r}

#species presence

head(horn1)

plot(msia.worldclim_raster) #predictors

horn.d = bioclim(msia.worldclim_raster, horn1[,c('long','lat')])
## bioclimatic model

ext = extent(99, 105, 1.2, 6.7)

backg = randomPoints(msia.worldclim_raster, n=1000,ext=ext, extf = 1.25)
#background/pseudo-absence data

e = evaluate(horn1, backg, horn.d, msia.worldclim_raster)
#presence, background, model, predictors

e

plot(e, 'ROC')

```

